---
title: "Optimal number of OID loops"
author: "Ally Hume"
date: "30 May 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

In this analysis we explore the optimal number of OID loops to execute within a 60 hour experiment.  The options explored are:

| Number of loops |  Duration of each OID input (hours) | 
|-----------------|-----------------------------|
| 1 | 60 | 
| 2 | 30 |
| 3 | 20 |
| 4 | 15 | 
| 5 | 12 |
| 6 | 10 |
| 8 | 4x8 and 4x7 |
| 10 | 6 | 
| 12 | 5 |
| 15 | 4 | 
| 20 | 3 | 


```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(knitr)

trueValues = data.frame( param=c("alpha1","Vm1","h1","Km1","d1","alpha2","d2","Kf","Kb"),
                         trueValue=c(3.3120e-05,0.0414,    1.89,   0.0368, 0.03,   0.01,   0.0100, 0.002, 0.0023) )

load_experiment_data <- function(filename) {
  
  wholeFile <- readLines(filename)
  iterationData <- wholeFile[grep('HOUR',wholeFile)]
  iterationData <- iterationData[grep('PARAM_FIT|REL_CONF',iterationData)]
  con <- textConnection(iterationData)
  data <- read.table(con)
  close(con)
  
  colnames(data) <- c("key", "hour", "type", "param", "value")
  
  data <- data %>% mutate( runName = gsub('.dat','',filename))
  numLoops <- filename
  numLoops <- gsub('CadBane',     '#1@',   numLoops)
  numLoops <- gsub('-2_loops',    '#2@',    numLoops)
  numLoops <- gsub('-3_loops',    '#3@',    numLoops)
  numLoops <- gsub('-4_loops',    '#4@',    numLoops)
  numLoops <- gsub('-5_loops',    '#5@',    numLoops)
  numLoops <- gsub('-6_loops',    '#6@',    numLoops)
  numLoops <- gsub('-8_loops',    '#8@',    numLoops)
  numLoops <- gsub('-10_loops',   '#10@',   numLoops)
  numLoops <- gsub('-12_loops',   '#12@',   numLoops)
  numLoops <- gsub('-15_loops',   '#15@',   numLoops)
  numLoops <- gsub('-20_loops',   '#20@',   numLoops)

  data <- data %>% mutate( numLoops = gsub('.*#','',numLoops))
  data <- data %>% mutate( numLoops = gsub('@.*','',numLoops))

  return(data)
}
```

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# Get the data files.
files <- c(list.files(path='../data',pattern='Gungi.*dat',full.names = TRUE),
           list.files(path='../data',pattern='BB8.*dat',full.names = TRUE),
           list.files(path='../data',pattern='CadBane.*dat',full.names = TRUE))

# Read in each file, convert to a data frame and merge into one dataframe
data <- lapply(files, load_experiment_data)
data <- Reduce(function(...) merge(..., all=T), data)

# Join data with true values
data <- left_join(data,trueValues)

data$numLoops <- factor(data$numLoops, levels = c('1','2','3','4','5','6','8','10','12','15','20'))
data$hour <- factor(data$hour)
```

## Number of runs

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'alpha1')
d <- d %>% group_by(numLoops) %>% summarise(n=n())
kable(d)
```

## Fitted parameter values after final hour 

The final fitted parameters for each method after 60 hours of experiment.

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 ) %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + facet_wrap( ~ param, scales = "free_y") + scale_y_log10() +
  geom_hline(aes(yintercept=trueValue), color='red')
```

The y-axis are poorly labelled so here we plot each parameter individually so we get better control over the axis labelling:

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'alpha1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10() +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("alpha1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'd1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=c(0.02,0.03,0.04)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("d1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'h1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(1.2,2.6,0.2)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("h1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'Km1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0.01,0.05,0.01)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("Km1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'Vm1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0.02,0.10,0.005)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("Vm1: Final fitted values")
```

## Relative confidence intervals after the final hour
  
Relative confidence intervals for each parameter by experiment. 

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0) %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + facet_wrap( ~ param, scales = "free_y") + scale_y_log10() 
```

The y-axis are poorly labelled so here we plot each parameter individually so we get better control over the axis labelling:

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0 & param == 'alpha1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=c(100,1000,10000,100000)) +
  ggtitle("alpha1: Final relative confidence")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0 & param == 'd1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0,100,10)) +
  ggtitle("alpha1: Final relative confidence")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0 & param == 'h1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0,100,10)) +
  ggtitle("h1: Final relative confidence")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0 & param == 'Km1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0,100,10)) +
  ggtitle("Km1: Final relative confidence")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0 & param == 'Vm1') %>% 
     select(numLoops, param, trueValue, value) 

ggplot(data=d, aes(x=numLoops,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0,100,10)) +
  ggtitle("Vm1: Final relative confidence")
```


## Normalised squared error over all parameters except alpha

Dividing the estimated parameter value by the true parameter value, summing the squared errors and then dividing by the number of data points gives the following graphs for each parameter (ignoring alpha1):

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param != 'alpha1') %>% 
     select(numLoops, param, trueValue, value) %>%
     mutate(normValue=value/trueValue) %>%
     mutate(squaredError=(normValue-value)^2) %>% 
     group_by(param,numLoops) %>%
     summarize(error=sum(squaredError)/n())

ggplot(data=d, aes(x=numLoops,y=error,group=1)) + 
  geom_line() + facet_wrap( ~ param, scales = "free_y") 
```

Combining these into a single graph (just using addition) gives:

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param != 'alpha1') %>% 
     select(numLoops, param, trueValue, value) %>%
     mutate(normValue=value/trueValue) %>%
     mutate(squaredError=(normValue-value)^2) %>% 
     group_by(numLoops) %>%
     summarize(error=sum(squaredError)/n())

ggplot(data=d, aes(x=numLoops,y=error,group=1)) + geom_line()  
```


## GitHub

The github repository containing the data and the processing files is:

* `https://github.com/csynbiosys/OID_In_Silico_Experiments`

The `README.md` file in that repository also contains references to the git repositories 
and versions of the code used to produce the data.




