---
title: "Comparison of means of creating the experiment input"
author: "Ally Hume"
date: "30 May 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Comparing various means of creating inputs for 60 hours of gal1 experiment:

Several different approaches to designing the experiment input where tried. These are described in the table below:

| Experiment name  | Description | Git branch |
|------------------|-------------|-----------------|
| OID (O)           | Uses OID to design the full input (60 min steps)    | Experiment-CadBane |
| Random (R)        | Randomly chooses a new input level (60 min steps) | Experiment-Ventress |
| Designed 13 (D13) | 13 input steps rising from minimum to maximum and back down again | Experiment-MazKanata |
| Designed 61 (D61) | 61 input steps rising from minimum to maximum and back down again | Experiment-Biggs |
| Pulse (P)         | 10 minute maximum input pulse every hour, otherwise zero galactose | Experiment-Lobot |
| Step 60 (S60)     | Alternate maximum and minimum galactose input for 6 hour periods | Experiment-Hondo |
| Step 1 (S1)       | Alternate maximum and minimum galactose input for 1 hour periods | Experiment-Phasma |

In each run the initial conditions are randomly chosen in the range 0.1 x truth to 10 x truth. The random distribution is an even distribution in the log scale, so half the distribution is in range [truth .. 10 x truth] and half in the range [0.1 x truth .. truth].

## Examples of input

The following graphs show example inputs used for the various experiments. For the OID and random experiment
the input is different for each run.  For the other experiments the same input is used for each run.


```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(knitr)

foo <- read.csv('../data/exampleInputs/CadBaneInputExample.dat',col.names=c('t','u'),header=FALSE)
ggplot(data=foo,aes(x=t,y=u)) + geom_line() + ggtitle("Example OID input")

foo <- read.csv('../data/exampleInputs/VentreeInputExample.dat',col.names=c('t','u'),header=FALSE)
ggplot(data=foo,aes(x=t,y=u)) + geom_line() + ggtitle("Example random input")

foo <- read.csv('../data/exampleInputs/BiggsInputExample.dat',col.names=c('t','u'),header=FALSE)
ggplot(data=foo,aes(x=t,y=u)) + geom_line() + ggtitle("D61 input")

foo <- read.csv('../data/exampleInputs/MazKanataInputExample.dat',col.names=c('t','u'),header=FALSE)
ggplot(data=foo,aes(x=t,y=u)) + geom_line() + ggtitle("D13 input")

foo <- read.csv('../data/exampleInputs/LobotInputExample.dat',col.names=c('t','u'),header=FALSE)
ggplot(data=foo,aes(x=t,y=u)) + geom_line() + ggtitle("Pulse input")

foo <- read.csv('../data/exampleInputs/HondoInputExample.dat',col.names=c('t','u'),header=FALSE)
ggplot(data=foo,aes(x=t,y=u)) + geom_line() + ggtitle("Step 60 input")

foo <- read.csv('../data/exampleInputs/PhasmaInputExample.dat',col.names=c('t','u'),header=FALSE)
ggplot(data=foo,aes(x=t,y=u)) + geom_line() + ggtitle("Step 1 input")

```


```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

trueValues = data.frame( param=c("alpha1","Vm1","h1","Km1","d1","alpha2","d2","Kf","Kb"),
                         trueValue=c(3.3120e-05,0.0414,    1.89,   0.0368, 0.03,   0.01,   0.0100, 0.002, 0.0023) )

load_experiment_data <- function(filename) {
  
  wholeFile <- readLines(filename)
  iterationData <- wholeFile[grep('HOUR',wholeFile)]
  iterationData <- iterationData[grep('PARAM_FIT|REL_CONF',iterationData)]
  con <- textConnection(iterationData)
  data <- read.table(con)
  close(con)
  
  colnames(data) <- c("key", "hour", "type", "param", "value")
  
  data <- data %>% mutate( runName = gsub('.dat','',filename))
  exptName <- filename
  exptName <- gsub('.*CadBane.*',   'O',    exptName)
  exptName <- gsub('.*Ventress.*',  'R',    exptName)
  exptName <- gsub('.*Biggs.*',     'D61',  exptName)
  exptName <- gsub('.*Lobot.*',     'P',  exptName)
  exptName <- gsub('.*MazKanata.*', 'D13',    exptName)
  exptName <- gsub('.*Phasma.*',    'S1',   exptName)
  exptName <- gsub('.*Hondo.*',     'S6',   exptName)
  
  filename <- gsub('.*CadBane.*',      'OID 1 Loop (CadBane)', filename)
  filename <- gsub('.*Ventress.*',     'Random (Ventress)',     filename)
  filename <- gsub('.*Biggs.*',        'Designed 61 (Biggs)',   filename)
  filename <- gsub('.*Lobot.*',        'Designed 13 (Lobot)',   filename)
  filename <- gsub('.*MazKanata.*',    'Pulse  (MaxKanata)',    filename)
  filename <- gsub('.*Phasma.*',       'Steps 1 hour (Phasma)', filename)
  filename <- gsub('.*Hondo.*',        'Steps 6 hour (Hondo)',  filename)
  data <- data %>% mutate( exptName = exptName)
  data <- data %>% mutate( exptLabel = filename)
  
  return(data)
}
```

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# Get the data files.
files <- c(list.files(path='../data',pattern='CadBane.*dat',   full.names = TRUE),
           list.files(path='../data',pattern='Ventress.*dat',  full.names = TRUE),
           list.files(path='../data',pattern='Biggs.*dat',     full.names = TRUE),
           list.files(path='../data',pattern='Lobot.*dat',     full.names = TRUE),
           list.files(path='../data',pattern='MazKanata.*dat', full.names = TRUE),
           list.files(path='../data',pattern='Phasma.*dat',    full.names = TRUE),
           list.files(path='../data',pattern='Hondo.*dat',     full.names = TRUE))

# Read in each file, convert to a data frame and merge into one dataframe
data <- lapply(files, load_experiment_data)
data <- Reduce(function(...) merge(..., all=T), data)

# Join data with true values
data <- left_join(data,trueValues)

# Specify the order of the experiment names
data$exptName <- factor(data$exptName, levels = c('O','R','D13','D61','P','S6','S1'))
data$hour <- factor(data$hour)
```

## Number of runs

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'alpha1')
d <- d %>% group_by(exptName,exptLabel) %>% summarise(n=n())
kable(d)
```

## Fitted parameter values after final hour 

The final fitted parameters for each method after 60 hours of experiment.

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 ) %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + facet_wrap( ~ param, scales = "free_y") + scale_y_log10() +
  geom_hline(aes(yintercept=trueValue), color='red')
```

The y-axis are poorly labelled so here we plot each parameter individually so we get better 
control over the axis labelling:

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'alpha1') %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10() +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("alpha1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'd1') %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0.01,0.10,0.01)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("d1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'h1') %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0.4,4,0.5)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("h1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'Km1') %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0.00,0.2,0.02)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("Km1: Final fitted values")
```

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param == 'Vm1') %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + scale_y_log10(breaks=seq(0.02,0.10,0.01)) +
  geom_hline(aes(yintercept=trueValue), color='red') +
  ggtitle("Vm1: Final fitted values")
```


Focusing on just OID and Random to zoom in a little more:

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & exptName %in% c('O','R')) %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + facet_wrap( ~ param, scales = "free_y") +
  geom_hline(aes(yintercept=trueValue), color='red')
```



## Convergence of OID method

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & exptName == 'O' ) %>% 
     select(hour, param, trueValue, value) 

ggplot(data=d, aes(x=hour,y=value)) + 
  geom_boxplot(alpha=0) + facet_wrap( ~ param, scales = "free_y") + scale_y_log10() +
  geom_hline(aes(yintercept=trueValue), color='red')
```

## Relative confidence intervals after the final hour
  
Relative confidence intervals for each parameter by experiment.  We have to be a little careful about these numbers because sometimes they are not produced and sometimes they are zero (which seems incorrect).  The data used to produce these is:

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0)
d <- d %>% group_by(exptName,exptLabel,param) %>% summarise(n=n())
kable(d)
```

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0) %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + facet_wrap( ~ param, scales = "free_y") + scale_y_log10() 
```

Focusing on just OID to zoom in a little more:

```{r echo=FALSE}
d <- data %>% filter(type == 'REL_CONF' & hour == 60 & value > 0 & exptName == 'O') %>% 
     select(exptName, param, trueValue, value) 

ggplot(data=d, aes(x=exptName,y=value)) + 
  geom_boxplot(alpha=0) + facet_wrap( ~ param, scales = "free_y")
```

## Normalised squared error over all parameters except alpha

Dividing the estimated parameter value by the true parameter value, summing the squared errors and then dividing by the number of data points gives the following graphs for each parameter (ignoring alpha1):

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param != 'alpha1') %>% 
     select(exptName, param, trueValue, value) %>%
     mutate(normValue=value/trueValue) %>%
     mutate(squaredError=(normValue-value)^2) %>% 
     group_by(param,exptName) %>%
     summarize(error=sum(squaredError)/n())

ggplot(data=d, aes(x=exptName,y=error,group=1)) + 
  geom_line() + facet_wrap( ~ param, scales = "free_y") 
```

Combining these into a single graph (just using addition) gives:

```{r echo=FALSE}
d <- data %>% filter(type == 'PARAM_FIT' & hour == 60 & param != 'alpha1') %>% 
     select(exptName, param, trueValue, value) %>%
     mutate(normValue=value/trueValue) %>%
     mutate(squaredError=(normValue-value)^2) %>% 
     group_by(exptName) %>%
     summarize(error=sum(squaredError)/n())

ggplot(data=d, aes(x=exptName,y=error,group=1)) + geom_line()  
```


## Code and data

The github repository containing the data and the processing files is:

* `https://github.com/csynbiosys/OID_In_Silico_Experiments`

The `README.md` file in that repository also contains references to the git repositories 
and versions of the code used to produce the data.



